---
title: John Hopkins University -- Data Science Specialization -- Regression Models
  Course -- Appendix to Solution of Project 1
author: "Dr. Guy Cohen"
date: "August 19, 2015"
output: pdf_document
---

# Effect of Transmission Type on Fuel Economy in 1973-74 Cars

## Appendix

### Exploratory Data Analysis

We first convert the factor variables to factor type.
```{r,fig.width=20,fig.height=20}
data(mtcars)
mtcars2 <- mtcars
mtcars2[,c("cyl","vs","am","gear","carb")] <-
    as.data.frame(lapply(mtcars2[,c("cyl","vs","am","gear","carb")],as.factor))
```

Second, we plot boxplots of continuous variables and barplots for categorical ones.  

```{r,fig.height=4,echo=FALSE}
par(mfrow=c(2,6),mar=c(4,4,2,1))
boxplot(mtcars2$mpg, ylab = "Miles/(US) gallon", xlab ="", main = "Fig. 1")
barplot(table(mtcars2$cyl), ylab = "Count", xlab = "Cylinder Num.", main = "Fig. 2")
boxplot(mtcars2$disp, ylab = "Displacement (cu.in.)", xlab ="", main = "Fig. 3")
boxplot(mtcars2$hp, ylab = "Gross Horsepower (hp)", xlab ="", main = "Fig. 4")
boxplot(mtcars2$drat, ylab = "Rear Axle Ratio", xlab ="", main = "Fig. 5")
boxplot(mtcars2$wt, ylab = "Weight (1000 lb)", xlab ="", main = "Fig. 6")
boxplot(mtcars2$qsec, ylab = "Quarter Mile Time (s)", xlab ="", main = "Fig. 7")
barplot(table(mtcars2$vs), ylab = "Count", xlab = "V/S", main = "Fig. 8")
barplot(table(mtcars2$am), ylab = "Count", xlab = "Transmission", main = "Fig. 9")
barplot(table(mtcars2$gear), ylab = "Count", xlab = "Fwd Gears", main = "Fig. 10")
barplot(table(mtcars2$carb), ylab = "Count", xlab = "Carburetors", main = "Fig. 11")
```

We plot a pairs plot to see how the variables relate to each other.
```{r, fig.width=20, fig.height=20, cache=TRUE}
library(ggplot2)
library(GGally)
ggpairs(mtcars2, lower = list(continuous = "smooth"), params = c(method = "loess"), title= "Fig. 12")
```

### Detailed Solution

We plot the adjusted R-squared for models of various number of predictors found using forward selection.
```{r, fig.height=2.5}
library(leaps)
library(ggplot2)
regfit.fwd=regsubsets(mpg ~ ., data = mtcars2, nvmax = 12, method="forward")
qplot(x,y,data = data.frame(x=1:12,y=summary(regfit.fwd)$adjr2), 
      geom = c("point","smooth"), xlab = "Number of Predictors", 
      ylab = "Adjusted R-Squared", main = "Fig. 13")
```

We fit a linear model using the parameters found and plot the residual plot.
```{r}
fit <- lm(mpg ~ I(cyl == "6") + hp + wt + I(am == "1"), data = mtcars2)
par(mfrow = c(2,2))
plot(fit)
```

